---
title: "Reproducible documents"
author: "Victoria Bøttker"
format: html
editor: visual
---

```{r setup}
#| message: false
#| warning: false
library(tidyverse)
post_meal_data <- read_csv2(here::here("data/post-meal-insulin.csv"))
```

## About me

-   Victoria Bøttker

-   Stud.med, Researh Year Student

-   Aarhus University

-   Aarhus University hospital, The department of Gynecology and Obstetrics

I am a 25-year-old **medical** student doing a research year at the department of Gynecology and obstetrics. The research is about threatened *preterm birth*.

## Simple code

```{r}
3*3
```

## Testing for Git 

The weather is medium. Not sunny, not rainy - but cloudy.

## Showing the data

```{r}
post_meal_data

```

```{r}
glimpse(post_meal_data)
```

## Plot one continous variable

```{r}
#| fig-cap: "Distribution of BMI."
#| label: fig-bmi-histo
ggplot(post_meal_data, aes(x = BMI)) + 
  geom_histogram()
```

## Exercise: discrete plots 

```{r}
#| fig-cap: "Group status after a meal."
#| label: fig-group-bar
ggplot(post_meal_data, aes(x = Group)) + 
  geom_bar()

```

@fig-group-bar

```{r}
#| fig-cap: "Distribution of glykemi."
#| label: fig-glykemi-bar
ggplot(post_meal_data, aes(x = glykemi)) + 
  geom_bar()
```

@fig-glykemi-bar

## Plotting two discrete variables 

```{r}
ggplot(post_meal_data, aes(x = Group, fill = as.character(glykemi)))+
  geom_bar(position = position_dodge())
```

## Side by side plots 

```{r}
ggplot(post_meal_data, aes(x = BMI)) + 
  geom_histogram()
```

```{r}
ggplot(post_meal_data, aes(x = glykemi)) + 
  geom_bar()
```

```{r side-by-side-figs}
#| label: fig-bmi-glycemia
#| fig-cap: "BMI and glycemia, side by side."
#| fig-subcap:
#| - "Distribution of BMI." 
#| - "Number of those with glycemia."
#| layout-ncol: 2
ggplot(post_meal_data, aes(x = BMI)) + 
  geom_histogram()

ggplot(post_meal_data, aes(x = glykemi)) + 
  geom_bar()

```
